### 2019-07-25 1_introduction.tex
More work on shrinking text.

Benchmarks in Task Bench are based on the observation that many applications can be modeled
as a set of \emph{tasks}, or coarse-grain units of work, with
dependencies between tasks representing the communication and synchronization required for
distributed execution. A benchmark consists of a \emph{task graph} with tasks for
each point in an iteration space. Tasks are connected by dependencies
as determined by a \emph{dependence relation}. This design permits a
very concise description of a wide variety of patterns relevant to
scientific computing and large scale data analysis: trivial parallelism, halo exchanges (such as
seen in structured and unstructured mesh codes), sweeps (such as used
in the discrete ordinates method of radiation simulation), FFTs, trees
(for divide and conquer algorithms), and so on. The tasks at each
point in the task graph can be configured to execute kernels with a
variety of computational properties, including compute- and
memory-bound operations of varying durations. The dependencies between
tasks can be configured to carry communication payloads of varying size, permitting
the design of communication-bound cases. Finally, multiple
(potentially heterogeneous) task graphs can be executed concurrently
to introduce task parallelism into the workload. Together, these
design elements permit the exploration of a large space of application
behaviors.

@@

Using Task Bench we were able to quickly incorporate 15 very different parallel and distributed
runtime systems.  By running all systems on common benchmarks we were able to quantify phenomena
that have never before been measured.
Most strikingly, the performance of the various systems we examine varies by more than five orders
of magnitude, with very popular, widely used systems at both ends of the spectrum!  Clearly
the slower systems have ``good enough'' performance for some class of applications, while presumably
providing advantages in programmer productivity.

@@

We conduct a comprehensive comparative study of Task Bench
implementations in 15 different programming systems:
Chapel \cite{Chapel15}, Charm++ \cite{Charmpp93}, Dask \cite{Dask15}, MPI \cite{MPI}, MPI+X (OpenMP and CUDA),
OmpSs \cite{OmpSs11}, OpenMP \cite{OpenMPSpec40},
PaRSEC \cite{PARSEC13}, Realm \cite{Realm14}, Regent \cite{Regent15},
Spark \cite{Spark10}, StarPU \cite{StarPU11},
Swift/T \cite{Wozniak13}, TensorFlow \cite{TensorFlow15}, and
X10 \cite{X1005}. Using METG(50\%), we determine the baseline
runtime overhead of each system along with overheads associated with
increasing numbers of and/or variations in the pattern of dependencies. We
also explore each system's ability to scale, hide communication
overhead, and mitigate load imbalance.

The paper is organized as follows: Section~\ref{sec:task-bench}
describes the design of Task Bench, Section~\ref{sec:implementation}
discusses the implementation of Task Bench in 15 programming systems,
Section~\ref{sec:metg} defines the METG(50\%) metric and its
relationship to quantities of interest to application developers,
Section~\ref{sec:evaluation} provides a comprehensive evaluation of
Task Bench on up to 256 Haswell nodes of the Cori
supercomputer~\cite{Cori}, Section~\ref{sec:related-work} relates this
work to previous efforts, and Section~\ref{sec:conclusion} concludes.

>>>>

Benchmarks in Task Bench are based on the observation that many applications can be modeled
as a set of \emph{tasks}, or coarse-grain units of work, with
dependencies between tasks representing the communication and synchronization required for
distributed execution. A benchmark consists of a \emph{task graph} with tasks for
each point in an \emph{iteration space}, and dependencies
determined by a \emph{dependence relation}. This design permits a
concise description of a wide variety of patterns relevant to
scientific computing and large scale data analysis: trivial parallelism, halo exchanges (such as
seen in structured and unstructured mesh codes), sweeps (such as used
in the discrete ordinates method of radiation simulation), FFTs, trees
(for divide and conquer algorithms), and so on. The tasks at each
point in the task graph can be configured to execute kernels with a
variety of computational properties, including compute- and
memory-bound operations of varying durations. The dependencies between
tasks can be configured to carry communication payloads of varying size, permitting
the design of communication-bound cases. Finally, multiple
(potentially heterogeneous) task graphs can be executed concurrently
to introduce task parallelism into the workload. Together, these
design elements permit the exploration of a large space of application
behaviors.

@@

Using Task Bench we were able to quickly incorporate 15 very different parallel and distributed
runtime systems.  By running all systems on common benchmarks we were able to quantify phenomena
that have never before been measured.
Most strikingly, the performance of the systems we examine varies by more than five orders
of magnitude, with very popular, widely used systems at both ends of the spectrum!  Clearly
the slower systems have ``good enough'' performance for some class of applications, while presumably
providing advantages in programmer productivity.

@@

We conduct a comprehensive comparative study of Task Bench
implementations in 15 different programming systems:
Chapel \cite{Chapel15}, Charm++ \cite{Charmpp93}, Dask \cite{Dask15}, MPI \cite{MPI}, MPI+X (OpenMP and CUDA),
OmpSs \cite{OmpSs11}, OpenMP \cite{OpenMPSpec40},
PaRSEC \cite{PARSEC13}, Realm \cite{Realm14}, Regent \cite{Regent15},
Spark \cite{Spark10}, StarPU \cite{StarPU11},
Swift/T \cite{Wozniak13}, TensorFlow \cite{TensorFlow15}, and
X10 \cite{X1005}. Using METG, we determine the baseline
runtime overhead of each system along with overheads associated with
increasing numbers of and/or variations in the pattern of dependencies. We
also explore each system's ability to scale, hide communication
overhead, and mitigate load imbalance.

The paper is organized as follows: Section~\ref{sec:task-bench}
describes the design of Task Bench, Section~\ref{sec:implementation}
discusses the implementation of Task Bench in 15 programming systems,
Section~\ref{sec:metg} defines the METG metric and its
relationship to quantities of interest to application developers,
Section~\ref{sec:evaluation} provides a comprehensive evaluation of
Task Bench on up to 256 Haswell nodes of the Cori
supercomputer~\cite{Cori}, Section~\ref{sec:related-work} relates this
work to previous efforts, and Section~\ref{sec:conclusion} concludes.

### 2019-07-25 2_task_bench.tex
More work on shrinking text.

Task graphs are a combination of an iteration space (with a task for
each point in the space) with a dependence relation.
For simplicity, but without loss of generality, the iteration space in
Task Bench is constrained to be 2-dimensional, with time along
the vertical axis and parallel tasks along the
horizontal. Tasks can depend only on tasks from the immediately
preceding time step. Figure~\ref{fig:task-graphs} shows a number of sample task
graphs that can be implemented with Task Bench. Note that layout is
significant---in particular, column $i$ represents all tasks
with index $i$ in the iteration space over all the time
steps. Generally speaking each column will be
assigned to execute on a different processor core.

@@

Despite its generality, Task Bench is easy to implement, making it
tractable to develop a suite of high-quality implementations. The central aspects of Task Bench, such as generating
task graphs and enumerating dependencies, are encapsulated in a core
library that is shared among all the Task Bench implementations. The
core library also includes implementations of the kernels, ensuring
that the kernels are identical in all systems, eliminating a potential
source of performance disparity that can be a pitfall for
implementations of mini-apps. Finally, the Task Bench core library manages the
parsing of input parameters and the display of output results,
ensuring that all implementations behave in a uniform manner and can
be scripted consistently. Because much of the functionality needed for
a Task Bench implementation is in the core library, implementations of
Task Bench are small: Our 15 Task Bench implementations range from 88
to 1500 lines, with several hundred lines being
typical. Listings~\ref{lst:compute-kernel} and \ref{lst:code-sample}
show excerpts from the Task Bench core, and the Dask implementation,
respectively. Only the code in Listing~\ref{lst:code-sample} is
implemented for each system, minimizing the work required for each additional system.

@@

Task Bench provides two main kernels that can be called from tasks,
which are compute- and memory-bound, respectively. The compute-bound
kernel executes a tight loop and is hand-written using AVX2 FMA
intrinsics. The memory-bound kernel performs sequential reads and
writes over an array, and is again hand-optimized with AVX2
intrinsics. The duration of both kernels can be configured by setting
the number of iterations to execute; we use this ability to simulate
the effects of varying application problem sizes. The memory-bound
kernel is carefully written to keep the working set size constant as
the number of iterations decreases, to avoid unwanted speedups due to
cache effects.

>>>>

Task graphs are a combination of an iteration space (with a task for
each point in the space) with a dependence relation.
For simplicity, but without loss of generality, the iteration space in
Task Bench is constrained to be 2-dimensional, with time along
the vertical axis and parallel tasks along the
horizontal. Tasks may depend only on tasks from the immediately
preceding time step. Figure~\ref{fig:task-graphs} shows a number of sample task
graphs that can be implemented with Task Bench. Note that layout is
significant---in particular, column $i$ represents all tasks
with index $i$ in the iteration space over all the time
steps. Generally speaking each column will be
assigned to execute on a different processor core.

@@

Despite its generality, Task Bench is easy to implement, making it
tractable to develop a suite of high-quality implementations. The central aspects of Task Bench, such as generating
task graphs and enumerating dependencies, are encapsulated in a core
library that is shared among all the Task Bench implementations. The
core library also includes implementations of the kernels, ensuring
that the kernels are identical in all systems, eliminating a potential
source of performance disparity that can be a pitfall for
implementations of mini-apps. Finally, the core library manages
parsing input parameters and displaying results,
ensuring that all implementations behave uniformly and can
be scripted consistently. Because much of the functionality needed for
a Task Bench implementation is in the core library, implementations of
Task Bench are small: Our 15 Task Bench implementations range from 88
to 1500 lines, with several hundred lines being
typical. Listings~\ref{lst:compute-kernel} and \ref{lst:code-sample}
show excerpts from the Task Bench core, and the Dask implementation,
respectively. Only the code in Listing~\ref{lst:code-sample} is
implemented for each system, minimizing the work required for each additional system.

@@

Task Bench provides two main kernels that can be called from tasks:
compute- and memory-bound. The compute-bound
kernel executes a tight loop and is hand-written using AVX2 FMA
intrinsics. The memory-bound kernel performs sequential reads and
writes over an array, again with AVX2
intrinsics. The duration of both kernels can be configured by setting
the number of iterations to execute; we use this ability to simulate
the effects of varying application problem sizes. The memory-bound
kernel is carefully written to keep the working set size constant as
the number of iterations decreases, to avoid unwanted speedups due to
cache effects.

### 2019-07-25 4_metg.tex
More work on shrinking text.

In Figure~\ref{fig:efficiency-mpi},
efficiency starts at 100\%. Initially task granularity
shrinks with minimal change in efficiency. However as task
granularity shrinks further, efficiency drops more rapidly, leading
to a vertical asymptote as runtime overhead dominates useful work.

@@

The METG metric has another useful property. Because METG is measured ``in place'' (i.e.,
without changing the number of nodes or cores available to the
application), METG isolates effects that are
due to shrinking problem size from effects that are due to
increased communication and other resource issues as
progressively larger portions of the machine are used. Section~\ref{sec:evaluation} contains a comprehensive evaluation of
METG including how it changes with node count for a wide variety of
programming systems.

>>>>

In Figure~\ref{fig:efficiency-mpi},
efficiency starts at 100\%. Initially task granularity
shrinks with minimal change in efficiency. As tasks shrink further, efficiency drops more rapidly, leading
to a vertical asymptote as overhead dominates useful work.

@@

The METG metric has another useful property. Because METG is measured ``in place'' (i.e.,
without changing the number of nodes or cores available to the
application), METG isolates effects
due to shrinking problem size from effects due to
increased communication and other resource issues as
progressively larger portions of the machine are used.
% Elliott: Do I need this sentence at all? The evaluation section is literally immediately after this.
%% Section~\ref{sec:evaluation} contains a comprehensive evaluation of
%% METG including how it changes with node count.
